---
title: "Final Project Presentation"
author: "Zachary Goodsell"
date: "12/11/2019"
output:
  ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
```

## Original Data

* Alarm Date
    + Month / Day / Year Format
    + Dates from 4/1/2012 to 8/15/2019

- Incident Types
    + Incident Type Number Followed by a Description
    + 111 - Building fire
    + 131 - Passenger vehicle fire
    + 164 - Outside mailbox fire

## Original Data Cont.

- Total Loss
    + Total Loss of the Property in a Fire
    + Min - 0
    + Max - 1,245,000

- Pre-Incident Value
    + Value of the Property Before an Incident
    + Min - 0
    + Max - 40,000,000

```{r}
firedata <- read.csv("PropertyLoss.csv")

#creat the month collum
firedata$Month <-  months(as.Date(firedata$Alm_Date, "%m/%d/%Y"))

#Turn month into the season
firedata <- transform(firedata, Season= ifelse(firedata$Month == "January", "Winter",
                  ifelse(firedata$Month == "February", "Winter", 
                  ifelse(firedata$Month == "March", "Spring",
                  ifelse(firedata$Month == "April", "Spring",
                  ifelse(firedata$Month == "May", "Spring",
                  ifelse(firedata$Month == "June", "Summer",
                  ifelse(firedata$Month == "July", "Summer",
                  ifelse(firedata$Month == "August", "Summer",
                  ifelse(firedata$Month == "September", "Fall",
                  ifelse(firedata$Month == "October", "Fall",
                  ifelse(firedata$Month == "November", "Fall",
                  ifelse(firedata$Month == "December", "Winter","NULL")))))))))))))
```

## Modifying the Data

```{r, echo = TRUE}
#parse out the incident numbers
firedata$Incident.Num <- strtoi(substr(firedata$Inci_type,0,3))

#collum with the incident type
firedata <- transform(firedata,Incident.Type = 
                   ifelse(Incident.Num < "124", "Structure Fire", 
                   ifelse(Incident.Num < "139", "Vehicle Fire",
                   ifelse(Incident.Num < "165", "Outside Fire",
                   "Not a Fire"))))

#Create the percent loss
firedata$PercentLoss <- (as.numeric(firedata$Pre_Inci_Value)
                         / as.numeric(firedata$Total_Loss))

#filter out non fires
firedata <- filter(firedata, Incident.Type != "Not a Fire")
```

## Modified Data

* Season
    + Spring - March, April, May
    + Summer - June, July, August
    + Fall - September, October, November
    + Winter - December, January, Febuary

* Incident Type
    + Structure Fire
    + Vehicle Fire
    + Outside Fire
    
* Percent Loss
    + Min - 0%
    + Max - 80.2%
```{r, include=F}
head(firedata[,c(11,13,14)])
```

## Anova Tests

Analysis of Variance (ANOVA) is a statistical method used to test differences between two or more means.

```{r, echo = TRUE}
#Anova test of total loss based on season
aovSeason = aov(PercentLoss ~ Season, data = firedata)
summary(aovSeason)
```

## Anova Tests Cont.

```{r, echo = TRUE}
#Anova test of total loss based on Incident.Type
aovI = aov(PercentLoss ~ Incident.Type, data = firedata)
summary(aovI)
```

## Summary of Percent Loss Based on Incident Type

```{r}
#Percent loss summary of each season based on incident type
firedata %>%
  group_by(Incident.Type) %>%
  summarize(Md = median(PercentLoss), Mean = mean(PercentLoss), SD = sd(PercentLoss), N = n())

```

## Box Plot of All the Months

The box plot is log scaled for the percent loss to see changes better

```{r}
#Winter Subset
winter <- firedata %>%
  filter(Season == "Winter") %>%
  group_by(Season) %>%
  select(Season, Incident.Type,Total_Loss,PercentLoss)

#Spring Subset
spring <- firedata %>%
  filter(Season == "Spring") %>%
  group_by(Season) %>%
  select(Season, Incident.Type,Total_Loss,PercentLoss)

#Summer Subset
summer <- firedata %>%
  filter(Season == "Summer") %>%
  group_by(Season) %>%
  select(Season, Incident.Type,Total_Loss,PercentLoss)

#Fall Subset
fall <- firedata %>%
  filter(Season == "Fall") %>%
  group_by(Season) %>%
  select(Season, Incident.Type,Total_Loss,PercentLoss)

#Box plot of all the subsets
boxplot(log(winter$PercentLoss),log(spring$PercentLoss),log(summer$PercentLoss),log(fall$PercentLoss),
        names = c("Winter","Spring","Summer","Fall"), xlab = "Season", ylab = "Log Total Loss")
```

## Summary of Percent Loss Based on Season

We see that the mean in the fall is lower then the mean in spring

```{r}
#Percent loss summary of each season
firedata %>%
  group_by(Season) %>%
  summarize(Md = median(PercentLoss), Mean = mean(PercentLoss), SD = sd(PercentLoss), N = n())
```

## T Test

```{r}
#T test, spring and fall 
t.test(spring$PercentLoss,fall$PercentLoss)
```


```{r, include= F}
#T test, winter and spring 
t.test(winter$PercentLoss,spring$PercentLoss)

#T test, winter and summer 
t.test(winter$PercentLoss,summer$PercentLoss)

#T test, winter and fall 
t.test(winter$PercentLoss,fall$PercentLoss)

#T test, spring and summer 
t.test(spring$PercentLoss,summer$PercentLoss)

#T test, spring and fall 
t.test(spring$PercentLoss,fall$PercentLoss)

#T test, summer and fall 
t.test(summer$PercentLoss,fall$PercentLoss)
```

# Question?
