---
title: "Math 275 Lab 1 Intro to Data"
author: "Zachary Goodsell"
date: "9/24/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("http://www.openintro.org/stat/data/cdc.R")

```

## Question 1
There are 20,000 cases and 9 variables for each of these cases. Data types - genhlth:categorical, exerany:binary, hlthplan:binary,smoke100:binary, height:continuous,weight:continuous,wtdesire:continuous, age:continuous,gender:binary
```{r cdc1}
names(cdc)
```


## Question 2
The interquartile range of height is 6. The interquartile range of age is 26. The gender frequency is 47.8% for males and 52.1% for females. The exerany frequency is 25.4% for 0 and 74.6% for 1. 0 is for no monthly exercise and 1 is for monthly exercise. There are 9569 males in the sample. For participants with excellent health the percent is 23.3%.
```{r cdc2}
summary(cdc$height)
summary(cdc$age)
iqr_height=summary(cdc$height)[5]-summary(cdc$height)[2]
iqr_height
#Interquartile range of height is 6
iqr_age=summary(cdc$age)[5]-summary(cdc$age)[2]
iqr_age
#Interquartile range of age is 26
gender=table(cdc$gender)
genderFreq = gender/20000
genderFreq
barplot(genderFreq)
exerany=table(cdc$exerany)
exeranyFreq = exerany/20000
exeranyFreq
barplot(exeranyFreq)
table(cdc$gender)
numMales = table(cdc$gender)[1]
numMales
table(cdc$genhlth)/20000
table(cdc$genhlth)[1]/20000
```
## Question 3
The mosaic shows the difference between long term smoking in men and women. We see that more men smoke 100 or more cigarettes in there lifetime when compared to woman.
```{r cdc3}
table(cdc$gender,cdc$smoke100)
mosaicplot(table(cdc$gender,cdc$smoke100),main="Mosaic Plot of Gender vs. smoke100")
title(ylab="Person Has Smoked More Than 100 Cigarettes in Lifetime",
      xlab="Persons Gender")
```

## Question 4
```{r cdc4}
under23_and_smoke = subset(cdc, smoke100==1 & age<23)
head(under23_and_smoke)
```

## Question 5
```{r cdc5, echo=FALSE}
bmi <- (cdc$weight / cdc$height^2) * 703
boxplot(bmi ~ cdc$genhlth)
```

The box plot above shows the relationship between the persons BMI and overall health.

```{r cdc6}

boxplot(bmi ~ cdc$exerany,
        main="Relationship Between BMI and Monthly Exercise",xlab="Whether or Not Person Has Exercised in Past Month",ylab="Persons BMI")
```

I chose monthly exercise because those who don't exercise tend to have a high BMI.The box plot shows that BMI is slightly lower in those who exercise.

## Own Questions
### Question 1
The relationship shows how close the peoples weights are to their desired weights. a perfect 1/1 slop would represent everyone being at there desired weight.  However, it is seen that the points aren't all at that slope.
```{r cdc7}
plot(cdc$weight, cdc$wtdesire,
     xlab="Weight",ylab="Desired Weight",main="Plot of Persons Actual Weights vs. Desired Weights")
```

### Question 2
```{r cdc8}
wdiff = abs(cdc$wtdesire-cdc$weight)
head(wdiff)
```

### Question 3
wdiff is quatitative data. If it is 0 that means that the persons actual weight and desired weight are the same. If wdiff isn't 0 that is fow far the person is from there desired weight.

### Question 4
The histogram is right skewed. The mean of 10 shows that 50% of the people are with in 10 pounds of there desired weight. Further more, the histogram shows that roughly 75% of the people are within 25 pounds of there desired weight.
```{r cdc9}
summary(wdiff)
hist(wdiff)
wdiff_under100=subset(wdiff, wdiff<=100)
summary(wdiff_under100)
hist(wdiff_under100)
```

### Question 5
The plot shows that men are closer to there desired weight. The women have a bigger spread.
```{r cdc10}
cdc_under100=subset(cdc,
                   abs(cdc$wtdesire-cdc$weight)<=100)
cdc_under100_male=subset(cdc_under100, gender=='m')
cdc_under100_female=subset(cdc_under100, gender=='f')
summary(abs(cdc_under100_male$wtdesire-cdc_under100_male$weight))
sd(abs(cdc_under100_male$wtdesire-cdc_under100_male$weight))
summary(abs(cdc_under100_female$wtdesire-cdc_under100_female$weight))
sd(abs(cdc_under100_female$wtdesire-cdc_under100_female$weight))
boxplot(abs(cdc_under100$wtdesire-cdc_under100$weight) ~ cdc_under100$gender, 
        main="Difference Between Actual and Desired Weight, by Gender",
        xlab="Gender",ylab="Difference Between Actual and Desired Weight")
```

### Question 6
70.76% of peoples weight fall within one SD of the mean weight
```{r cdc11}
weight_mean = mean(cdc$weight)
weight_sd = sd(cdc$weight)
weight_mean
weight_sd
sum(cdc$weight>=(weight_mean-weight_sd) & cdc$weight<=(weight_mean+weight_sd))/nrow(cdc)

```
