---
title: "MATH 275 Final Exam Part II"
author: "Zachary Goodsell"
date: "12/10/2019"
output:  
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
library(knitr)
library(kableExtra)
sleepAp <- read_excel("OSASdataFinalExam.xlsx")
```

## Question 1

Create a new variable in R for the waist-to-height ratio (waist circumference / height) from the dataset and define it as WtoH for statistical analysis -

```{r}
#Create the waist-to-height ratio collum
sleepAp$WtoH <- as.numeric(sleepAp$WAIST) / as.numeric(sleepAp$HEIGHT)
head(sleepAp$WtoH)
```

Create new categories in R for the severity of OSAS, based on the above information, name them as mildOSAS, moderateOSAS, and severeOSAS - 

```{r}
#collum with the OSAS severity
sleepAp <- transform(sleepAp, severity  = 
                   ifelse(as.numeric(AHI) < as.numeric("5"), "Control",
                   ifelse(as.numeric(AHI) < as.numeric("15"), "mildOSAS",
                   ifelse(as.numeric(AHI) < as.numeric("30"), "moderateOSAS",
                   "severeOSAS"))))
```

## Question 2 

There were 509 cases in the study. The mean age was 47.86 (with sd= 12.5, min=18, max=78). Of these cases, 437 (86%) were diagnosed with OSAS and 72 (14%) were not. Two hundred and nine (41%) of the patients had severe, 124 (24%) had moderate, and 104 (20%) had mild OSAS. 

```{r, include=FALSE}
length(sleepAp$age)

summary(sleepAp$age)

sd(sleepAp$age)

table(sleepAp$severity)

(437)/(437 + 72)

(72)/(437 + 72)

(209)/(437 + 72)

(124)/(437 + 72)

(104)/(437 + 72)

```

The mean age of patients with OSAS (mean ± sd = 49.27±12.01) was significantly higher than those without OSAS (mean ± sd = 39.35±12.09) (p-value 4.444e-09, Welch t-test).

```{r, include=FALSE}
osas <- filter(sleepAp, severity != "Control")

NOosas <- filter(sleepAp, severity == "Control")

summary(osas$age)

sd(osas$age)

summary(NOosas$age)

sd(NOosas$age)

t.test(osas$age,NOosas$age)

```

## Question 3

One hundred fifty (29.4%) of the participants were female and 359 were male. Of the patients with OSAS, 114 were female (27%) and 323 were male (74%). There were 36 females and 36 male participants who were not diagnosed with OSAS. 

```{r, include=FALSE}
table(sleepAp$sex)

150/(150+359)

table(osas$sex)

114 / (114 + 323 )

323 / (114 + 323 )

table(NOosas$sex)


```

```{r, echo=FALSE}

osasTable <- data.frame("Sex" = c("Female","Male"), "OSAS Patients" = c("114","323"), "Control Group" = c("36","36"))
kable(osasTable)

```

Table 1. Number of female and male cases in patient and control groups.

There was an association between sex and incidence of OSAS (p-value 6.765e-05, Chi-square test). It means that, the incidence of OSAS among males (90%) is higher than that in females (78%) (p-value 0.000154, two-sample proportion test).

```{r, include=FALSE}
#collum with the OSAS or not
sleepAp <- transform(sleepAp, present  = 
                   ifelse(as.numeric(AHI) < as.numeric("5"), "no",
                   ifelse(as.numeric(AHI) < as.numeric("15"), "yes",
                   ifelse(as.numeric(AHI) < as.numeric("30"), "yes",
                   "yes"))))

chisq.test(sleepAp$present,sleepAp$sex)


prop.test(c(323,114),c(355,146))



```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
