---
title: "Lab 4"
author: "Zachary Goodsell"
date: "10/20/2019"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("prettydoc")
```

## Part A
In the SienaMath-height.csv file, µ = 68 inches and σ = 3.763863 inches.

```{r a, echo=FALSE}
hgt=read.csv("SienaMath-height.csv")
mean(hgt$height)
sd(hgt$heigh)
```

## Part B

The sample of the SienaMath-height.csv file is -

```{r b, echo=FALSE}
s=sample(hgt$height,3)
s
```

The mean of this sample is -

```{r b2, echo=FALSE}
mean(s)
```

## Part C

Here the final value is the mean of all the means of the data.

```{r c, echo=FALSE}
MeanXbar=numeric(10)
for (i in 1:10){smp=sample(hgt$height,3);MeanXbar[i]=mean(smp)}

MeanXbar

mean(MeanXbar)
```

## Part D

```{r d1, echo=FALSE}
MeanXbar=numeric(100)
for (i in 1:100){smp=sample(hgt$height,3);MeanXbar[i]=mean(smp)}
hist(MeanXbar,col="orange", xlab="Mean of Samples",main="")
```

The mean of the sample means is -

```{r d2, echo=FALSE}
mean(MeanXbar)
```

The standard deviation of the sample means is -

```{r d3, echo=FALSE}
sd(MeanXbar)
```

The approximate mean of the sample means is -

```{r d4, echo=FALSE}
mean(hgt$height)
```


The approximate standard deviation of the sample means is -

```{r d5, echo=FALSE}
sd(hgt$height)/sqrt(3)
```

These values compare such that µ is the mean of the histogram and √σ/n is one standard deviation.

## Part E

```{r e1, echo=FALSE}
MeanXbar=numeric(5984)
for (i in 1:5984){smp=sample(hgt$height,3);MeanXbar[i]=mean(smp)}
hist(MeanXbar,col="green", xlab="Mean of Samples",main="")
```

The mean of the sample means is -

```{r e2, echo=FALSE}
mean(MeanXbar)
```

The standard deviation of the sample means is -

```{r e3, echo=FALSE}
sd(MeanXbar)
```

The approximate mean of the sample means is -

```{r e4, echo=FALSE}
mean(hgt$height)
```


The approximate standard deviation of the sample means is -

```{r e5, echo=FALSE}
sd(hgt$height)/sqrt(3)
```

## Part F

### Sample Size = 1
```{r f1, echo=FALSE}
xbar=numeric(1000)
for (i in 1:1000){x=runif(n=1,min=0,max=1);xbar[i]=mean(x)}
hist(xbar,col="yellow",main="sample size=1",xlab="sample mean Xbar")
```

The mean is - 

```{r f2, echo=FALSE}
mean(xbar)
```

The standard deviation is -

```{r f3, echo=FALSE}
sd(xbar)
```

### Sample Size = 30
```{r f4, echo=FALSE}
xbar=numeric(1000)
for (i in 1:1000){x=runif(n=30,min=0,max=1);xbar[i]=mean(x)}
hist(xbar,col="red",main="sample size=30",xlab="sample mean Xbar")
```

The mean is - 

```{r f5, echo=FALSE}
mean(xbar)
```

The standard deviation is -

```{r f6, echo=FALSE}
sd(xbar)
```

### Sample Size = 1000
```{r f7, echo=FALSE}
xbar=numeric(1000)
for (i in 1:1000){x=runif(n=1000,min=0,max=1);xbar[i]=mean(x)}
hist(xbar,col="green",main="sample size=1000",xlab="sample mean Xbar")
```

The mean is - 

```{r f8, echo=FALSE}
mean(xbar)
```

The standard deviation is -

```{r f9, echo=FALSE}
sd(xbar)
```
