---
title: "Math275_Homework3"
author: "Zachary Goodsell"
date: "10/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 3.6

Sum of 1 probability - 
```{r sumProb1, echo = FALSE}
#The first die has to be a 1 therefore the next one has to be a 0
1/6*0/6
```

Sum of 5 probability - 
```{r sumProb2, echo = FALSE}
#the first die has to be a 1-4 therefore there is only one other possiblilty the othe die has to be
4/6*1/6
```

Sum of 12 probability - 
```{r sumProb3, echo = FALSE}
#the first die has to be a 6 therefore there is only one other possiblilty the othe die has to be - 6
1/6*1/6
```

## Question 3.9

a- Having two randomly selected students from your class that both earn A's are independent because student 1's grade doesn't affect students 2 grade.

b- Having both you and your class study partner earn an A in the course is neither because if the two students get A's that means they studied well together.

c- No because the two events may have no affect on each other.  

## Question 3.16

a- No, because those without excellent health also have health coverage.

b - 23.29%

c- 
```{r health1, echo = FALSE}
0.2099/0.8738
#= 24.02%
```

d- 
```{r health2, echo = FALSE}
0.0230/0.1262 
#= 18.26%
```

e- No because when compared to people of the same health coverage status it shows that the approximate percent of excellent health is the same.

## Question 3.20

```{r prob1, echo = FALSE}
(.99*0.03)/(.99*.03 + .02 * .97)

```

## Question 3.30

a- 
Expected winnings -
```{r prob2, echo = FALSE}

#probabily of winning nothing 
x0 = 26/52

#probabily of winning 5 
x5 = 13/52

#probabily of winning 10 
x10 = 12/52

#probabily of winning 20 
x20 = 1/52

expected_winnings = x0 * 0 + x5 * 5 + x10 * 10 + x20 * 20

expected_winnings


```
Standard Deviation -
```{r prob3, echo = FALSE}
total_mon = 13 * 5 + 12 * 10 + 1 *20

sd2 = ((0-total_mon/52)^2 * x0) + ((5-total_mon/52)^2 * x5) + ((10-total_mon/52)^2 * x10) + ((20-total_mon/52)^2 * x20)
  
sqrt(sd2)
  
```

b- The maximum I would bet is $3.94 because with that bet I should always be able to break even.


## Question 3.38

a- 
The distribution of total personal income is that 69.3% of the incomes are between $15k and $65k

b- 
```{r prob4, echo = FALSE}
income_perc <- c(0.022, 0.047, 0.158, 0.183, 0.212, 0.139, 0.058, 0.084, 0.097)
male_perc <- 0.59
female_perc <- 0.41

sum(income_perc[1:5])

```

c- Here we have to assume that all females are evenly distributed across all income ranges.

```{r prob5, echo = FALSE}
sum(income_perc[1:5]) * female_perc

```

d- My assumption that all females are evenly distributed was wrong because my percent for females that make less than $50,000 per year was 25% not 72% that was given.