---
title: "Math 275 Homework 2"
author: "Kursad Tosun"
date: "September 24, 2019"
output:
  html_document:
    highlight: pygments
    keep_md: no
    theme: cerulean
  pdf_document:
    fig_caption: no
  word_document: default
---

```{r setup,  include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1.** Write your answer in here.

* item 1
* item 2

**2.**  Write your answer in here. You can use the following table, or read Lab-RMarkdown for better table options.

Table Header | Second Header  
------------- | -------------  
Table Cell | Cell 2   
Cell 3 | Cell 4    

**3.**  Write your answer in here. The following R code chunk will display both the R code and results.   

```{r, warning = FALSE, fig.align='center', fig.cap='Histogram of hospital stays', out.width = "50%"}
library(readxl)
heart <- read_excel("Heart_attack_stays.xlsx")
head(heart)
hist(heart$Stay, breaks=30, xlab ='Stay (days)', ylab='Number of Female Patients', main='')
summary(heart)
```

Since the variables Sex and Hospital are categorical, we need encode them as a factor.  
```{r, warning = FALSE, fig.align='center', fig.cap='Histogram of hospital stays', out.width = "50%"}
heart$Sex=factor(heart$Sex)
heart$Hospital=factor(heart$Hospital)
summary(heart)
```

***
# Your answer should be short, clear and descriptive.

Let's consider a hypothetical homework question. 

**How many patients stayed in the NYU hospital?**

Your answer should include

1. a related R code and output (do not include unnecessary output), 
2. answer of the question.

**Solution.**
```{r, warning = FALSE}
summary(heart$Hospital)
```

Sixty-five patients were stayed in the NYU hospital.



**Note.** Your answer, R code and output should be short, clear and descriptive. If we talk about this specific example, your solution should not contain any results other than `Hospital`. Do **NOT include**  all the summary statistics in your answer. 

***


